Many claims of group polarization are plausibly false due to inappropriate statistics

### 2022-12-01 GroupPolarizationStatmod/paper/ms.tex
updated paper model section

We assume that participant $i$ has latent opinion at time $t$ that is drawn from
a group-level normal distribution,
\begin{equation}
  o_{i,t} \sim p(o; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
  \label{eq:opinionDistribution}
\end{equation}
\noindent 
This distribution has mean $\mu_t$ and standard deviation $\sigma_t$.
This form is chosen to
reflect the common statistical analysis used by group polarization studies,
not because it is statistically appropriate---in fact it is not, which is what
we aim to show in this paper.

@@

\begin{table}[h]
  \begin{tabular}{cp{5.0in}} \toprule
   Symbol & Description  \\ \midrule  
   $t \in \{\mathrm{pre}, \mathrm{post}\}$ & Time, either \emph{pre}- or
                                           \emph{post}-discussion \\
   $\mu$  &   Latent mean, assumed constant for simple consensus \\ 
   $\sigma_t$ & Latent standard deviation at time $t$ \\
   $p(o;\mu,\sigma_t)$ & Normal latent distribution of opinions with mean
                         $\mu$ and variance $\sigma_t^2$ \\
   $k \in \{0,\ldots,K\} $ & Threshold index ($k$) between 0 and total number of
         ordinal bins, $K$ \\
   $b_k$ & Opinion bin $k$; note there is no $k=0$ bin. \\
   $\theta_{k-1},~\theta_k$  & Lower, upper threshold value for binning latent opinions into
     bin $b_k$ \\
   $p(o \in b_k; \mu, \sigma_t)$ & Ordinal density of binned opinions in bin $b_k$ 
            (Equation~\ref{eq:binFrequency}) \\
   $\langle o_t \rangle$  &   Observed mean opinion calculated from ordinal data at time $t$ \\
   \bottomrule
  \end{tabular} 
\end{table}

>>>>

We begin by developing a formal model of ordinal measurement of latent continuous
opinions (model variables and parameters are given in Table~\ref{tab:modelVariables}).
First, we assume that participant $i$ has latent opinion at time $t$ that is drawn from
a group-level normal distribution,
\begin{equation}
  o_{i,t} \sim p(o; \mu_t, \sigma_t) = \mathcal{N}(\mu_t, \sigma_t).
  \label{eq:opinionDistribution}
\end{equation}
\noindent 
This distribution has mean $\mu_t$ and standard deviation $\sigma_t$.
This form is chosen to
reflect the common statistical analysis used by group polarization studies,
not because it is statistically appropriate---in fact it is not, which is what
we aim to show in this paper.

@@

\begin{table}[h]
  \caption{Variables used in the formal model of group polarization from simple consensus.}
  \label{tab:modelVariables}
  \begin{tabular}{cp{5.0in}} \toprule
   Symbol & Description  \\ \midrule  
   $t \in \{\mathrm{pre}, \mathrm{post}\}$ & Time, either \emph{pre}- or
                                           \emph{post}-discussion \\
   $\mu$  &   Latent mean, assumed constant for simple consensus \\ 
   $\sigma_t$ & Latent standard deviation at time $t$ \\
   $p(o;\mu,\sigma_t)$ & Normal latent distribution of opinions with mean
                         $\mu$ and variance $\sigma_t^2$ \\
   $k \in \{0,\ldots,K\} $ & Threshold index ($k$) between 0 and total number of
         ordinal bins, $K$ \\
   $b_k$ & Opinion bin $k$; note there is no $k=0$ bin. \\
   $\theta_{k-1},~\theta_k$  & Lower, upper threshold value for binning latent opinions into
     bin $b_k$ \\
   $p(o \in b_k; \mu, \sigma_t)$ & Ordinal density of binned opinions in bin $b_k$ 
            (Equation~\ref{eq:binFrequency}) \\
   $\langle o_t \rangle$  &   Observed mean opinion calculated from ordinal data at time $t$ \\
   \bottomrule
  \end{tabular} 
\end{table}
